---
title: "Course Project 1"
author: "Pozhitkova"
date: '20 February 2018 '
output:
  html_document:
  keep_md: true
---

---

### Reproducible Research Course Project 1

```{r, results = "asis"}
pacman::p_load(plyr, dplyr, tidyr)
pacman::p_load(readr, haven, ggplot2)
select<-dplyr::select
```



```{r global_options, include=FALSE}
knitr::opts_chunk$set(echo = TRUE,cache = TRUE,  fig.path='figure/')
```

## Loading and preprocessing the data

Load the data

```{r load}
setwd("C:/!KRISTINA/!COURSERA/R_Reproducible Research/Course Project 1")
if(!file.exists('activity.csv')){
    unzip('activity.zip')
}
df <- read.csv('activity.csv')
```

## What is mean total number of steps taken per day?

1. Calculate the total number of steps taken per day

```{r totalstep}
steps_day <- df %>% group_by(date) %>% 
        summarise(total_steps_by_day=sum(steps, na.rm=TRUE))
```

2. Make a histogram of the total number of steps taken each day

```{r hist}
ggplot(steps_day,  aes(total_steps_by_day)) + geom_histogram() + xlab("total number of steps taken per day")
```

3. Calculate and report the mean and median of the total number of steps taken per day
```{r mean}
mean_steps <- round(mean(steps_day$total_steps_by_day),1)
median_steps <- median(steps_day$total_steps_by_day)
```

The mean = `r mean_steps` and median = `r median_steps` of the total number of steps taken per day.


## What is the average daily activity pattern?

1. Time series plot  of the 5-minute interval  and the average number of steps taken, averaged across all days 

```{r stepsinterval}
steps_interval <- df %>% group_by(interval) %>% 
        summarise(mean_steps_interval=mean(steps, na.rm=TRUE)) 

```

```{r plotmean}
ggplot(data=steps_interval, aes(x=interval, y=mean_steps_interval)) + geom_line()
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

```{r whichmax}
interval_max_steps<-(steps_interval%>%arrange(desc(mean_steps_interval)))$interval[1] 
```

The maximum number of steps in  interval = `r interval_max_steps` 

## Imputing missing values

1. Calculate and report the total number of missing values in the dataset 

```{r countNA}
countNA<-dim(df%>%filter(is.na(steps)))[1]
```

The total number of missing values in the dataset `r countNA`.

3. Create a new dataset that is equal to the original dataset but with the missing data filled in.

```{r imputation}
df_imp <- df%>% group_by(interval) %>% 
        mutate_at(vars(steps) , funs(ifelse(is.na(.),mean(.,na.rm=TRUE), .)))
```


4. Make a histogram of the total number of steps taken each day and Calculate and report the mean and median total number of steps taken per day.

```{r totelstepsimp}
steps_day_imp <- df_imp %>% group_by(date) %>% 
        summarise(total_steps_by_day=sum(steps, na.rm=TRUE))

ggplot(steps_day_imp,  aes(total_steps_by_day)) +   geom_histogram()
```


```{r meanimp}
mean_steps_imp <- round(mean(steps_day_imp$total_steps_by_day),1)
median_steps_imp <- median(steps_day_imp$total_steps_by_day)
```

After missing imputation (NA=mean by date)  values are higher the mean = `r mean_steps_imp` and median = `r median_steps_imp` of the total number of steps taken per day. for some days there were no values of the variable steps. Based on the missing values, the estimates were displaced


## Are there differences in activity patterns between weekdays and weekends?

1. Create a new factor variable in the dataset with two levels – “weekday” and “weekend” indicating whether a given date is a weekday or weekend day.

```{r computetype}
df_imp_type<-df_imp%>% mutate(type = weekdays(as.POSIXlt(date)))  %>% 
        mutate(type=ifelse(type=="ñóááîòà"|type=="âîñêðåñåíüå",'weekend', 'weekday'))%>%
        group_by(type, interval)%>%summarise(mean_steps_interval=mean(steps, na.rm=TRUE)) 
```

2. Make a panel plot containing a time series plot 

```{r plotcompare}
ggplot(data=df_imp_type, aes(x=interval, y=mean_steps_interval)) + geom_line() +facet_grid(type ~ .)
```



